Guidelines and Policy for the KIDS Strong Lensing Science Team using \SW

### 2016-05-23 doc/sw-publication-policy-kids.tex
KIDS Guidelines

\title{Guidelines and Policy for the Dark Energy Survey Strong Lensing Science Team using \SW}
\author{Aprajita Verma, Phil Marshall \& Anupreeta More}
\date{March 25, 2015}
\maketitle

@@

\begin{abstract} 
\noindent \SW was conceived as a strong lens
discovery {\it service} that harnesses the power of citizen scientists. The DES strong lensing group (hereafter DES survey science team or DES SST) can make use of \SW within the logistical constraints. In this document we discuss how the DES SST can become involved with \SW and the relationship
between the DES SST and the \SW collaboration. 
This document includes guidelines for proposing projects, data provision, data policy and for publications of \SW-enabled discoveries.
\end{abstract}

@@

In order to use the \SW interface, the DES Survey Science Team \textbf{must define and propose a project to the \SW PIs} who will discuss whether the project is suitable with the SST. The \SW PIs will then liaise further with the SST to implement the project in the \SW interface. Details are given in Section \ref{sec:project}.

@@

In this document, we will refer to the entirety of the DES strong lensing science group and the STRIDES team as the DES Survey Science Team (DES SST) to encompass all of the DES strong lensing community.

We now discuss the relationship between the \SW collaboration and the DES SST, who wish to use the \SW interface to investigate their data. The DES SSTs will liaise with the \SW science and development teams to ingest their survey data into the \SW interface. There may be some overlap between current \SW members and SSTs. In the case of DES, \SW PI Anupreeta More is a member of both the \SW collaboration and the DES Strong Lensing Science Group.

@@

If required, an expanded definition of the \SW membership is given in the document ``\SW Publication Policy: CFHT-LS''\footnote{available from \texttt{\url{https://github.com/drphilmarshall/SpaceWarps/blob/master/doc/sw-publication-policy-cfhtls.pdf}}}, however this will not be relevant for \SW-DES publications, see Section \ref{sec:papers} for details.

@@

The DES SST is expected to propose a project to the \SW collaboration that can be best or only be done using the power of citizen scientists. Typically, this involves visual inspection tasks that are too time consuming for the SST themselves to execute. As such, any science programme conducted through \SW should be complementary, not in competition, to the lens finding activities being carried out by the SST.

@@

\subsection{DES specific thoughts}
In the case of DES, with its long data release plan and large strong lens group, we request the SST to consider the following:
\begin{enumerate}
\item The \SW facility should be used for science projects that cannot be completed by the Strong Lens Group itself. 
\item Any \SW search should be complementary, not in competition, to the lens finding being conducted by the DES SST.
\item While \SW uses citizen science, its primary goal is \textbf{not outreach}, it is harnessing citizen power for visual classifications (and eventually modelling) to perform science.
\item A duplicate project is not encouraged unless there is sufficient scientific motivation and justification to do so.
\item \SW requires volunteers/citizens to give up their personal time to conduct the experiment, this should not be taken lightly, the crowd are motivated by contributing to real science not just doing things for the sake of them. This is the reason the suitability of the project for citizen scientists should be carefully considered. The \SW PIs will be able to discuss any planned \SW-DES projects with the DES SST. It is important to bear this in mind for long-term surveys like DES, if needed the crowd needs to be engaged and feel like they are contributing while the entire \SW-DES project runs.
\item The success of \SW-DES will rely on input from the DES SST working with the \SW collaboration, dedicated effort is required for this and the DES SST should identify team members who can commit time to the \SW-DES tasks for this before embarking on a project. The \SW PIs will help to establish the main efforts required, these will depend on the actual project put forward by the DES SST (e.g.\ image preparation, simulations, interacting with volunteers on \Talk).
\end{enumerate}

\textbf{As such, \SW should be considered to be an integral part of the DES SST lens finding strategy, rather than an alternative method.}

@@

The goal of the \SW PIs and the \SW collaboration is to enable the DES SST's science project in \sw and make it work. Using the automatic analysis pipeline (SWAPR), the \SW PIs we will send the DES SST lists of the candidates selected by the citizens (at some regular intervals or at the end of a project), we will then discuss strategies for further analysis based on our experience of previous projects. Some members of the \SW PIs  and team may become more involved in analysis steps such as grading of the \sw candidates or assist with any scientific investigations and follow-up of the candidates, at the DES SST's request.

\subsection{Previewing images and \Talk}
We are aware that there is some interest in the DES SST for using the \SW web interface to preview images for the SST themselves. This is not the function or purpose of the \SW interface or the Zooniverse. However,  it will be possible for DES SST members to view the images in parallel with the crowd through the public interface.  The DES SST members who wish to do this will need to register with the Zooniverse to classify. We can extract the classifications of SSTs only to provide an expert selected sample, if required. The images marked by any inspector (SST or citizen)  are automatically recorded in an individual's \SW/Zooniverse account. These can be accessed through the \SW interface and as a 'collection' in \Talk. This collection can be explored (and discussed if required) through the \Talk interface.

\SW\ \Talk is a key and integral part of \SW. The DES SST is strongly encouraged to interact with the volunteers on \Talk. This exchange allows the crowd to learn from the experts, this in turn has the very positive effect of motivating the citizens to contribute further. In addition it also results in better and informed classification when they return to that task. Thus, engagement of the SST is very important in log-term surveys such as DES.

\subsection{Possible \SW-DES survey strategies}

In our experience, a citizen science project like SW cannot function successfully without significant support and co-ordination of the SST. Thus, we recommend that all DES members with an interest in strong lensing are told about \SW and how it can complement their lens finding effort. Our goal is to enable your science. While we can't expect the entirety of the DES Strong Lensing Group or the STRIDES consortium to be behind \SW-DES, all should be made aware of the projects being undertaken and planned with \SW.

We also encourage the DES SST to consider the timing for when they wish to ingest the images into \SW so that it is in line with the interests of the DES SST and those of the DES consortium.

Possible projects for the DES SST could consider (just to get a flavour of what can be done)
\begin{itemize}
\item \textbf{New Data}: Inspection of areas (targeted or blind) that the SST has not had time to inspect, i.e. use the crowd to increase the speed at which potential lens candidates are discovered
\item \textbf{Comparison of Expert vs Citizen Classification}: DES SST and citizen's view images in parallel (targeted or blind)
\item \textbf{Completeness of Expert Classification}: Citizen's classify images already inspected by the DES SST (targeted or blind)
\item \textbf{Helping the robots}: Use citizens to weed out false positives from candidate samples generated by algorithm based searches
\item \textbf{Training sets}: Blind inspection to generate a large training set for  machine learning algorithms.
\end{itemize}

Note that the projects that are proposed to \SW need not be one of the above. Multiple projects could also be run in a mixture or phased set of experiments. Please do consider the DES timeline with reference to data release and the searches the DES SST want to make themselves.

If the DES SST wants to inspect targets within the consortium and is concerned about being scooped by the crowd, please be assured that there is little scope or motivation to do this. If the project is well defined, the goal of \SW-DES will be a complementary part of the DES SST lens finding strategy. By definition this means that \SW is working with the DES SST as it is designed to do. In addition, the images could be stripped of WCS info (described in Section \ref{sec:data}), meaning that no-one other than the SST can follow-up or catalogue the targets\footnote{We cannot exclude the possibility that individual sources might be serendipitously identified e.g. by recognisable morphologies.}.  The DES SST could also perform their inspection before the images are passed to \SW. Any lens candidates discovered by the SST would then be labelled as `known lens candidates' in \sw\,\Talk. This applies to candidates discovered through visual inspection or algorithm based searches.

\subsection{What to do next: Project Proposals}
Once the DES SST have narrowed down what search to be conducted \SW, please approach the \SW PIs (Phil Marshall, Anupreeta More, Aprajita Verma, collectively contacted via \href{mailto:spacewarpspi@googlegroups.com}{spacewarpspi@googlegroups.com}) with a short description of the project that includes the following information.

\begin{itemize}
\item Survey area
\item Bands that could be included
\item Total number of images to be viewed (approx.)
\item Targeted survey or blind field search\footnote{Targeted: inspection of a list of candidates; Blind: inspection of tiles of a survey area}
\item Brief outline of the overall goal(s) of that inspection, highlighting how this is complementary (not competing) with the goals of the DES lens finding group
\item Proprietary data issues 
\item Identify an initial core team who will lead and work closely with \SW
\end{itemize}

Note, the proposed science project(s) should harness citizen classification power to achieve the science goals.

The \SW PIs will then arrange a telecon with the DES SST to discuss the project further. This includes discussing the initial effort required and roles for the the SST team for expected tasks.

The original \SW papers, by definition, provide a complete description of the system, and its results from the CFHT-LS survey. The SST will be able
to read these papers before designing their own \SW project, and cite them as justification for some of their experimental decisions. As the \SW paper series get published, links to the papers will be given in this draft. Prior to publications, draft papers can also be shared on request by the \SW PIs (\href{mailto:spacewarpspi@googlegroups.com}{spacewarpspi@googlegroups.com})

@@

However, some surveys, including DES, come with their own proprietary access policy. It is the responsibility
of the DES SST to provide images to \SW in a way that is consistent
with their own data access rules. There are two things that SSTs can
do to in order to respect any proprietary period that they have imposed on
themselves:
\begin{enumerate}

@@

\item In {\bf preparing data} for the site, and uploading it onto the
Zooniverse servers. There is a reasonable amount of work in preparing data and images by the DES SST for delivery to \SW. The \SW team will provide assistance to the DES SST in data preparation. The benefits of using the \SW service will far outweigh the time needed to prepare data.  The \SW Science Team can advise on suitable Training
Subjects (images of simulations and false positives) and image display settings, and also on formatting the data
ready for ingestion and display. A mandatory element of the training subjects is the production of large samples of simulated lenses. \SW will provide resources to assist generating these simulations.

\item In {\bf reconfiguring the site itself} ready for the new data. The DES
SST will need to include new Spotter's Guide images and text,
modified tutorial content, and additional survey-specific site content;
the \SW Science Team can help with all of this.

@@

\item In {\bf running the \SW analysis pipeline} (SWAPR) which will be the \sw PIs responsibility. This will provide citizen selected candidates to be sent to the DES SST.

@@

\section{Resulting publications for \SW-DES }
\label{sec:series}

@@

Any \SW collaboration member, simply by virtue of their Zooniverse
registration, may write a Zooniverse
Letter\footnote{\texttt{http://letters.zooniverse.org}} describing their
investigation of any lens candidate they find in \SW. This, along with
posts in  \Talk, is the primary means by which we expect \SW-DES collaboration
members (citizens participating in the project) will communicate their findings to the rest of the
astronomical community. The investigation of any \SW images that are
provided without world coordinate system (WCS) information will be
necessarily limited but the \SW Zooniverse \Letters will contribute to the collective knowledge and understanding of the presented system(s). Zooniverse \Letters are citable \SW
objects, and in some cases will appear listed on ADS\footnote{To be confirmed.}.

@@

\section{Summary of Publication Guidelines}
\label{sec:publ}
Any \SW series and \SW enabled discovery papers should include
\begin{itemize}
\item citations to the \SW system papers, \SW I \& II
\item a small number of \SW authors proposed by the \SW PIs (see \ref{sec:papers}).
\item the collaboration should be acknowledged in the acknowledgement section with a link to the collaboration membership page (URL: to be confirmed when available).
\end{itemize}

>>>>

\title{Guidelines and Policy for the KIDS Strong Lensing Science Team using \SW}
\author{Aprajita Verma, Phil Marshall \& Anupreeta More}
\date{May 23, 2016}
\maketitle

@@

\begin{abstract} 
\noindent \SW was conceived as a strong lens discovery {\it service}
that harnesses the power of citizen scientists. The KIDS strong
lensing group (hereafter KIDS survey science team or KIDS SST) can
make use of \SW within the logistical constraints. In this document,
we discuss how the KIDS SST might become involved with \SW and the
relationship between the KIDS SST and the \SW collaboration.  This
document includes guidelines for proposing projects, data provision,
data policy and for publications of \SW-enabled discoveries.
\end{abstract}

@@

In order to use the \SW interface, the KIDS Survey Science Team
\textbf{must define and propose a project to the \SW PIs} who will
discuss whether the project is suitable with the KIDS SST. The \SW PIs will
then liaise further with the SST to implement the project in the \SW
interface. Details are given in Section \ref{sec:project}.

@@

In this document, we will refer to the entirety of the KIDS strong
lensing science group as the KIDS Survey Science Team (KIDS SST) to
encompass all of the KIDS strong lensing community.

We now discuss the relationship between the \SW collaboration and the
KIDS SST, who wish to use the \SW interface to investigate their
data. The KIDS SSTs will liaise with the \SW science and development
teams to ingest their survey data into the \SW interface. There may be
some overlap between current \SW members and SSTs.

In the case of KIDS, \SW PI Anupreeta More is a member of both the \SW
collaboration and the KIDS Strong Lensing Science Group.

@@

If required, an expanded definition of the \SW membership is given in
the document ``\SW Publication Policy: CFHT-LS''\footnote{available
  from
  \texttt{\url{https://github.com/drphilmarshall/SpaceWarps/blob/master/doc/sw-publication-policy-cfhtls.pdf}}},
however this will not be relevant for \SW-KIDS publications, see
Section \ref{sec:papers} for details.

@@

The KIDS SST is expected to propose a project to the \SW collaboration
that can be best or only be done using the power of citizen
scientists. Typically, this involves visual inspection tasks that are
too time consuming for the SST themselves to execute. As such, any
science programme conducted through \SW should be complementary, not
in competition, to the lens finding activities being carried out by
the SST.

@@

\subsection{KIDS specific thoughts}

\begin{enumerate}
\item The \SW facility should be used for science projects that cannot be completed by the Strong Lens Group itself. 
\item Any \SW search should be complementary, not in competition, to the lens finding being conducted by the KIDS SST.
\item While \SW uses citizen science, its primary goal is \textbf{not outreach}, it is harnessing citizen power for visual classifications (and eventually modelling) to perform science.
\item A duplicate project is not encouraged unless there is sufficient scientific motivation and justification to do so.
\item \SW requires volunteers/citizens to give up their personal time to conduct the experiment, this should not be taken lightly, the crowd are motivated by contributing to real science not just doing things for the sake of them. This is the reason the suitability of the project for citizen scientists should be carefully considered. The \SW PIs will be able to discuss any planned \SW-KIDS projects with the KIDS SST. The crowd needs to be engaged and feel like they are contributing while the entire \SW-KIDS project runs.
\item The success of \SW-KIDS will rely on input from the KIDS SST working with the \SW collaboration, dedicated effort is required for this and the KIDS SST should identify team members who can commit time to the \SW-KIDS tasks for this before embarking on a project. The \SW PIs will help to establish the main efforts required, these will depend on the actual project put forward by the KIDS SST (e.g.\ image preparation, simulations, interacting with volunteers on \Talk).
\end{enumerate}

\textbf{As such, \SW should be considered to be an integral/complementary part of the KIDS SST lens finding strategy, rather than an alternative method.}

@@

The goal of the \SW PIs and the \SW collaboration is to enable the
KIDS SST's science project in \sw and make it work. Using the
automatic analysis pipeline (SWAPR), the \SW PIs we will send the KIDS
SST lists of the candidates selected by the citizens (at some regular
intervals or at the end of a project), we will then discuss strategies
for further analysis based on our experience of previous
projects. Some members of the \SW PIs and team may become more
involved in analysis steps such as grading of the \sw candidates or
assist with any scientific investigations and follow-up of the
candidates, at the KIDS SST's request.

\subsection{Possible \SW-KIDS survey strategies}

In our experience, a citizen science project like SW cannot function
successfully without significant support and co-ordination of the
SST. Thus, we recommend that all KIDS members with an interest in
strong lensing are told about \SW and how it can complement their lens
finding effort. Our goal is to enable your science.

We also encourage the KIDS SST to consider the timing for when they
wish to ingest the images into \SW so that it is in line with the
interests of the KIDS SST and those of the KIDS consortium.

Possible projects for the KIDS SST could consider (just to get a flavour of what can be done)
\begin{itemize}
\item \textbf{New Data}: Inspection of areas (targeted or blind) that the SST has not had time to inspect, i.e. use the crowd to increase the speed at which potential lens candidates are discovered
\item \textbf{Comparison of Expert vs Citizen Classification}: KIDS SST and citizen's view images in parallel (targeted or blind)
\item \textbf{Completeness of Expert Classification}: Citizen's classify images already inspected by the KIDS SST (targeted or blind)
\item \textbf{Helping the robots}: Use citizens to weed out false positives from candidate samples generated by algorithm-based or machine learning searches
\item \textbf{Training sets}: Blind inspection to generate a large training set for machine learning algorithms.
\end{itemize}

Note that the projects that are proposed to \SW need not be one of the
above. Multiple projects could also be  run in a mixture or phased set
of experiments. Please do consider the KIDS timeline with reference to
data release and the searches the KIDS SST want to make themselves.

If the KIDS SST wants to inspect targets within the consortium and is
concerned about being scooped by the crowd, please be assured that
there is little scope or motivation to do this. If the project is well
defined, the goal of \SW-KIDS will be a complementary part of the KIDS
SST lens finding strategy. By definition this means that \SW is
working with the KIDS SST as it is designed to do. In addition, the
images could be stripped of WCS info (described in Section
\ref{sec:data}), meaning that no-one other than the SST can follow-up
or catalogue the targets\footnote{We cannot exclude the possibility
  that individual sources might be serendipitously identified e.g. by
  recognisable morphologies.}.  The KIDS SST could also perform their
inspection before the images are passed to \SW. Any lens candidates
discovered by the SST would then be labelled as `known lens
candidates' in \sw\,\Talk. This applies to candidates discovered
through visual inspection or algorithm based searches.

\subsection{What to do next: Project Proposals}
Once the KIDS SST have narrowed down what search to be conducted \SW,
please approach the \SW PIs (Phil Marshall, Anupreeta More, Aprajita
Verma, collectively contacted via
\href{mailto:spacewarpspi@googlegroups.com}{spacewarpspi@googlegroups.com})
with a short description of the project that includes the following
information.

\begin{itemize}
\item Survey area
\item Bands that could be included
\item Total number of images to be viewed (approx.)
\item Targeted survey or blind field search\footnote{Targeted: inspection of a list of candidates; Blind: inspection of tiles of a survey area}
\item Brief outline of the overall goal(s) of that inspection, highlighting how this is complementary (not competing) with the goals of the KIDS lens finding group
\item Proprietary data issues 
\item Identify an initial core team who will lead and work closely with \SW
\end{itemize}

The \SW PIs will then arrange a telecon with the KIDS SST to discuss
the project further. This includes discussing the initial effort
required and roles for the the SST team for expected tasks.

The original \SW papers, by definition, provide a complete description of the system, and its results from the CFHT-LS survey. The SST will be able
to read these papers before designing their own \SW project, and cite them as justification for some of their experimental decisions. 
See \href{https://github.com/drphilmarshall/SpaceWarps/raw/master/doc/sw-system-published.pdf}{Paper I} \& \href{https://github.com/drphilmarshall/SpaceWarps/raw/master/doc/sw-cfhtls-published.pdf}{Paper II}.

@@

However, some surveys, including KIDS, come with their own proprietary
access policy. It is the responsibility of the KIDS SST to provide
images to \SW in a way that is consistent with their own data access
rules. There are two things that SSTs can do to in order to respect
any proprietary period that they have imposed on themselves:
\begin{enumerate}

@@

\item In {\bf preparing data} for the site, and uploading it onto the
Zooniverse servers. There is a reasonable amount of work in preparing data and images by the KIDS SST for delivery to \SW. The \SW team will provide assistance to the KIDS SST in data preparation. The benefits of using the \SW service will far outweigh the time needed to prepare data.  The \SW Science Team can advise on suitable Training
Subjects (images of simulations and false positives) and image display settings, and also on formatting the data
ready for ingestion and display. A mandatory element of the training subjects is the production of large samples of simulated lenses. \SW will provide resources to assist generating these simulations.

\item In {\bf reconfiguring the site itself} ready for the new data. The KIDS
SST will need to include new Spotter's Guide images and text,
modified tutorial content, and additional survey-specific site content;
the \SW Science Team can help with all of this.

@@

\item In {\bf running the \SW analysis pipeline} (SWAPR) which will be the \sw PIs responsibility. This will provide citizen selected candidates to be sent to the KIDS SST.

@@

\section{Resulting publications for \SW-KIDS }
\label{sec:series}

@@

Any \SW collaboration member, simply by virtue of their Zooniverse
registration, may write a Zooniverse
Letter\footnote{\texttt{http://letters.zooniverse.org}} describing their
investigation of any lens candidate they find in \SW. This, along with
posts in  \Talk, is the primary means by which we expect \SW-KIDS collaboration
members (citizens participating in the project) will communicate their findings to the rest of the
astronomical community. The investigation of any \SW images that are
provided without world coordinate system (WCS) information will be
necessarily limited but the \SW Zooniverse \Letters will contribute to the collective knowledge and understanding of the presented system(s). Zooniverse \Letters are citable \SW
objects.

@@

\section{Summary of Publication Guidelines}
\label{sec:publ}
Any \SW series and \SW enabled discovery papers should include
\begin{itemize}
\item citations to the \SW system papers, \SW I \& II
\item a small number of \SW authors proposed by the \SW PIs (see \ref{sec:papers}).
\item the collaboration should be acknowledged in the acknowledgement section with a link to the collaboration membership page (e.g. \href{https://spacewarps.org/#/projects/CFHTLS/contributors}{List of Contributors to CFHT-LS}).
\end{itemize}
